Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-01 ms.tex
Updates from Overleaf

Many models of social learning assume little, or nothing, cognitive capabilities
of their agents. For example, a common modelling strategy compare the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions  ~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we need to
endow agents with a biologically plausible cognition capable of integrating
information from various sources in a payoff-maximizing way~\cite{Gershman2019}. 
% \cm{I don't
%   understand this next sentence, but think it can be cut with no loss of meaning:
%   In our model we do not impose any structure in randomness between behaviors, so
% our cognitive model is only sensitive to total randomness.} 
To do so, we let
agents see the payoffs to behaviors and use the softmax function to convert these
observations to a probability distribution from which they strategically select
their own behavior. %\cm{does this still make sense? and does it need a citation?}

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values and observed how frequently social learning evolved across
1000 trial simulations. We varied $u \in
\{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$;
and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  We
observed three outcome measures, with $\meansl$ as the primary measure: 
the average value of $s_i$ over all agents and trials. To support our conclusions
based on $\meansl$, we also measured $\meanpi / L$, the average net 
payoffs at simulation end across agents and trials, normalized by lifespan; 
and $\meanT / L$, or the mean
number of time steps to fixation normalized by lifespan. We analyze outcomes by
plotting $\meansl$ on the y-axis and environmental variation is on x-axis since we
theoretically expect that $\meansl$ will decrease monotonically from 1 to 0 as $u$
increases. Note that the hypothesis-testing concept of significance is meaningless
here because we could make any small outcome difference ``significant'' by running
more simulation trials.  We instead study the patterns of outcome variables over
systematically varied uncertainty factor values.

@@

        $\meanT / L$ & Mean number of generations to convergence & 20k / 8 max. \\
        \bottomrule
    \end{tabular}
\end{table}

>>>>

Many models of social learning assume little to no cognitive capabilities
of their agents. For example, a common modelling strategy compare the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions  ~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we need to
endow agents with a biologically plausible cognition capable of integrating
information from various sources in a payoff-maximizing way~\cite{Gershman2019}. 
% \cm{I don't
%   understand this next sentence, but think it can be cut with no loss of meaning:
%   In our model we do not impose any structure in randomness between behaviors, so
% our cognitive model is only sensitive to total randomness.} 
To do so, we let
agents see the payoffs to behaviors and use the softmax function to convert these
observations to a probability distribution from which they strategically select
their own behavior. %\cm{does this still make sense? and does it need a citation?}

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values and observed how frequently social learning evolved across
1000 trial simulations. We varied $u \in
\{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$;
and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  We
observed three outcome measures, with $\meansl$ as the primary measure: 
the average value of $s_i$ over all agents and trials. To support our conclusions
based on $\meansl$, we also measured $\meanpi / L$, the average net 
payoffs at simulation end across agents and trials, normalized by lifespan; 
and $\meanG$, the mean
number of generations to fixation. We analyze outcomes by
plotting $\meansl$ on the y-axis and environmental variation is on x-axis since we
theoretically expect that $\meansl$ will decrease monotonically from 1 to 0 as $u$
increases. Note that the hypothesis-testing concept of significance is meaningless
here because we could make any small outcome difference ``significant'' by running
more simulation trials.  We instead study the patterns of outcome variables over
systematically varied uncertainty factor values.

@@

        $\meanG$ & Mean number of generations to convergence & 20k / 8 max. \\
        \bottomrule
    \end{tabular}
\end{table}
